Computing solution space properties of combinatorial optimization problems via generic tensor networks

### 2022-08-29 paper/paper.tex
Updates from Overleaf

\section{Tensor networks}\label{sec:tn}
\begin{definition}[Tensor Network~\cite{Cirac2021, Orus2014}]
    A tensor network is a quadruple of $\mathcal{N} = (\Tensors, \Inputs, \Output, \Sigma)$ where $\Sigma$ is a the alphabet of labels, $\Tensors = (\Tensors_1, \ldots, \Tensors_m)$ is a vector of tensors, $\Inputs$ is a map from tensor to a vector of labels $\Inputs: \Tensors_j \mapsto \mathbf{l}_j$, $(\mathbf{l}_j)_i \in \Sigma, i=1,\ldots, \rm{rank}(\Tensors_j)$.
    $\Output$ is a vector of labels for specifying the output.
\end{definition}
A tensor network is composed of a collection of tensors, a collection of labels associated with tensor dimensions, and an indicator of which tensor is the output.
The operation to evaluate a tensor network is called \textit{contraction}, which is defined as a summation of tensor element products over the labels not appearing in the output tensor.
For example, the matrix multiplication is a special tensor network that can be represented as $C_{ik} = A_{ij}B_{jk}$, where $A, B$ and $C$ are matrices (two-dimensional tensors), $ik,ij$ and $ik$ are labels associated with them, and ``$=$'' is the indicator of $C$ being the output, while its contraction is defined as $C_{ik} = \sum_j A_{ij}B_{jk}$.
%can be viewed as a generalization of binary matrix multiplication to nary tensor contraction.
%Generalized Einstein's notation is often used to represent a tensor network, where we use a label in the subscript to represent a degree of freedom.
%One can enumerate these degrees of freedom and accumulate the product of tensor elements to the output tensor.
% Einstein's notation is often used to represent a tensor network, e.g.\ it represents the matrix multiplication between two matrices $A$ and $B$ as $C_{ik} = A_{ij}B_{jk}$,
% where we use a label in the subscript to represent a degree of freedom.
% One can enumerate these degrees of freedom and accumulate the product of tensor elements to the output tensor.
% In the standard Einstein's notation for tensor networks in physics, each index appears precisely twice, either both in input tensors (which will be summed over) or one in a input tensor and another in the output tensor.
% Hence a tensor network can be represented as a open simple graph,
% where a tensor is mapped to a vertex, a label shared by two input tensors is mapped to a normal edge and a label appears in the output tensor is mapped to an open edge.
% In this work, we adopt a more general form of tensor networks that do not restrict how many times a label can appear in tensors. 
The graphical representation of a tensor network is an open (meaning an edge can connect to external vertices) hypergraph, where an input tensor is mapped to a vertex and a label is mapped to a hyperedge that can connect an arbitrary number of vertices, while the labels appearing in the output tensor are open.
Our notation is a minor generalization of the standard tensor network notation used in physics as we do not restrict the number of times a label can appear in the tensors. 
While this generalized form is equivalent in representation power, it can have smaller contraction complexity as will be illustrated in \Cref{sec:tensorbad}.
% because one can easily translate a generalized tensor network to the standard notation by adding $\delta$ tensors, a high dimensional equivalence of the identity matrix.
%However, introducing $\delta$ tensors may significantly increase the complexity of the tensor contraction. We illustrate this subtle point in \Cref{sec:tensorbad}.

\begin{example}
$C_{ijk} = A_{jkm} B_{mil} V_{jm}$ is a tensor network that can be evaluated as $C_{ijk} = \sum_{ml}A_{jkm} B_{mil} V_{jm}$.
Its hypergraph representation is shown below, where we use different colors to represent different hyperedges.

@@

\section{Tensor network representation of independent sets} \label{sec:tnmap}
This section describes the reduction of the independent set problem to a tensor network contraction problem.
An alternative interpretation, perhaps more accessible to physicists, can be found in \Cref{sec:energymodel}, where we introduce the reduction from the energy model of hardcore lattice gases~\cite{Dyre2016, Fernandes2007}.
Let $G=(V,E)$ be an undirected graph with each vertex $v\in V$ being associated with a weight $w_v$.
An independent set $I \subseteq V$ is a set of vertices that for any vertex pair $u,v \in I$, we have $(u, v) \not\in E$.
To reduce the independent set problem on $G$ to a tensor network contraction, we first map each vertex $v\in V$ to a label $s_v \in \{0, 1\}$ of dimension $2$, where we use $0$ ($1$) to denote a vertex is absent (present) in $I$.
For each label $s_v$, we define a parameterized rank-one tensor indexed by it as
\begin{equation}
    W(x_v^{w_v})_{s_v} = \left(\begin{matrix}
        1 \\
        x_v^{w_v}
    \end{matrix}\right)_{s_v},\label{eq:vertextensor}
\end{equation}
where $x_v$ is a variable associated with $v$.
Tensor elements can be indexed by subscripts, e.g.\ $W(x_v^{w_v})_0=1$ is the first element associated with $s_v=0$ and $W(x_v^{w_v})_1=x_v^{w_v}$ is the second element associated with $s_v=1$.
Similarly, for each edge $(u, v) \in E$, we define a matrix $B$ indexed by $s_u$ and $s_v$ as
\begin{equation}
    \qquad \quad 
       B = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right), \label{eq:edgetensor}
\end{equation}
which encodes the independent set constraint since $B_{11} = 0$, meaning the two vertices cannot be both in the independent set.

The corresponding tensor network contraction can be defined as
\begin{equation}\label{eq:idp}
    P(G, \{x_1^{w_1}, x_{2}^{w_2}, \ldots,x_{|V|}^{w_{|V|}}\}) = \sum\limits_{s_1, s_2, \ldots, s_{|V|} = 0}^{1} \prod\limits_{v\in V} W(x_v^{w_v})_{s_v} \prod\limits_{(u,v) \in E} B_{s_u s_v},
\end{equation}
where the summation runs over all $2^{|V|}$ vertex configurations $\{s_1, s_{2}, \ldots,s_{|V|}\}$ and accumulates the product of tensor elements to the output $P$. 
% It is easy to see that an edge tensor element $B_{s_us_v}$ encodes the independent set constraint. It zeros out the product whenever both $u$ and $v$ are in the independent set.
A vertex tensor element $W(x_v^{w_v})_{s_v}$ contributes a multiplicative factor $x_v^{w_v}$ whenever $v$ is in the set.

@@

    \def\dx{5};
    \def\r{0.25cm}
    \foreach \x/\y/\e in {1.25/0/ab}
        \node[color=black,fill=black,dot=2*\r] at (\x+\dx,\y) (\e) {\scriptsize \color{white}{B}};
    \foreach \x/\y/\v in {0.3/0/a, 2.2/0/b}
        \node[color=black] at (\x+\dx,\y) (s\v) {$s_\v$};
    \foreach \x/\y/\v in {-0.5/0/a, 3.0/0.0/b}{
        \node[dot=\wr, color=black] at (\x+\dx,\y) (w\v) {};
        \node at (\x+\dx,\y+0.5) () {\scriptsize \color{black}{$W(x_\v^{w_\v})$}};
    }
    \draw [cyan,thick] (wa) -- (sa);
    \draw [cyan,thick] (sa) -- (ab);
    \draw [red,thick] (wb) -- (sb);
    \draw [red,thick] (sb) -- (ab);
\end{tikzpicture}}

In the graphical representation of the tensor network on the right panel,
we use a circle to represent a tensor, a hyperedge in cyan color to represent the degree of freedom $s_a$,
and a hyperedge in red color to represent the degree of freedom $s_b$.
Tensors sharing the same degree of freedom are connected by the same hyperedge.
The contraction of this tensor network has the following form:
\begin{equation}
    \begin{split}
    P({\rm K2},\{x_a^{w_a}, x_b^{w_b}\}) &=
    \left(\begin{matrix} 1 & x_a^{w_a} \end{matrix}\right)
    \left(\begin{matrix} 1 & 1\\ 1  & 0 \end{matrix}\right)
    \left(\begin{matrix} 1 \\ x_b^{w_b} \end{matrix}\right)\\
    &= 1 + x_a^{w_a} + x_b^{w_b}.
    \end{split}
\end{equation}
The resulting polynomial represents 3 different independent sets $\{\}$, $\{a\}$, and $\{b\}$, with weights being $0$, $w_a$, and $w_b$ respectively.
\end{example}

For a general graph, it is computationally inefficient to evaluate \Cref{eq:idp} by directly summing up the $2^{|V|}$ products.
A better approach to evaluate a tensor network is: to find a good pair-wise tensor contraction order as a binary tree and then contract two tensors at a time by this order.
Let us denote the line graph of the hypergraph representation of a tensor network as $G'$.
A contraction order of this tensor network corresponds to a tree decomposition of $G'$ and the largest intermediate tensor has a rank equal to the width of this tree decomposition~\cite{Markov2008}.
Therefore, an optimal (in terms of space complexity) contraction order corresponds to the tree decomposition of $G'$ with the smallest width (or the treewidth ${\rm tw}(G')$).
In the independent set problem, the treewidth of $G'$ is the same as the treewidth of the graph this tensor network is mapped from. 
In practice, it is difficult to find an optimal contraction order for large tensor networks because finding the treewidth is a well-known NP-hard problem.
However, it is easy to find a close-to-optimal contraction order within typically a few minutes with a heuristic algorithm~\cite{Kourtis2019, Kalachev2021}.
%The pairwise tensor contraction also makes it possible to utilize basic linear algebra subprograms (BLAS) functions to speed up the computation for certain tensor element types.
For large-scale applications, it is also possible to slice over certain degrees of freedom to reduce the space complexity, i.e.\
loop and accumulate over certain degrees of freedom so that one can have a smaller tensor network inside the loop due to the removal of these degrees of freedom.

\begin{example}\label{eg:tensorcontraction}
In this example, we map a 5-vertex graph (left) to a tensor network (right) and show how the contraction order reduces the time and space complexities.

\centerline{\begin{tikzpicture}[
dot/.style = {circle, fill, minimum size=#1,
            inner sep=0pt, outer sep=0pt},
dot/.default = 6pt  % size of the circle diameter 
                ]  
    \def\dx{0};
    \def\r{0.25cm}
    \filldraw[fill=black] (\dx,0) circle [radius=\r];
    \filldraw[fill=black] (\dx,1.5) circle [radius=\r];
    \filldraw[fill=black] (1.5+\dx,0) circle [radius=\r];
    \filldraw[fill=black] (1.5+\dx,1.5) circle [radius=\r];
    \filldraw[fill=black] (2.5+\dx,2.5) circle [radius=\r];
    \draw [black,thick] (\dx,0) -- (\dx,1.5);
    \draw [black,thick] (\dx,0) -- (1.5+\dx,0);
    \draw [black,thick] (\dx,1.5) -- (1.5+\dx,1.5);
    \draw [black,thick] (1.5+\dx,0) -- (1.5+\dx,1.5);
    \draw [black,thick] (1.5+\dx,0) -- (\dx,1.5);
    \draw [black,thick] (2.5+\dx,2.5) -- (1.5+\dx,1.5);
    \node[color=white] at (\dx,0) {a};
    \node[color=white] at (\dx,1.5) {b};
    \node[color=white] at (1.5+\dx,0) {c};
    \node[color=white] at (1.5+\dx,1.5) {d};
    \node[color=white] at (2.5+\dx,2.5) {e};
    \def\dx{5};
    \def\r{0.25cm}
    \foreach \x/\y/\e in {0.75/0/ac, 0/0.75/ab, 1.5/0.75/cd, 0.75/1.5/bd, 0.75/0.75/bc, 2/2/de}
        \node[color=white,fill=black,dot=2*\r] at (\x+\dx,\y) (\e) {\scriptsize B};
    \foreach \x/\y/\v in {0/0/a, 0/1.5/b, 1.5/0/c, 1.5/1.5/d, 2.5/2.5/e}
        \node[color=black] at (\x+\dx,\y) (\v) {$s_\v$};
    \foreach \x/\y/\v in {-0.5/-0.5/a, -0.5/2.0/b, 2.0/-0.5/c, 2.0/1.0/d, 3.0/2.0/e}
        \node[color=white,fill=black,dot=\r] at (\x+\dx,\y) (\v\v) {};
    \foreach \x/\y/\v in {-0.5/-0.5/a, -0.5/2.0/b, 2.0/-0.5/c, 2.0/1.0/d, 3.0/2.0/e}
        \node[color=black] at (\x+\dx+0.6,\y) {\scriptsize $W(x_\v^{w_\v})$};
    \draw [cyan,thick] (a) -- (aa);
    \draw [cyan,thick] (a) -- (ab);
    \draw [cyan,thick] (a) -- (ac);
    \draw [blue,thick] (b) -- (bb);
    \draw [blue,thick] (b) -- (ab);
    \draw [blue,thick] (b) -- (bc);
    \draw [blue,thick] (b) -- (bd);
    \draw [red,thick] (c) -- (cc);
    \draw [red,thick] (c) -- (ac);
    \draw [red,thick] (c) -- (bc);
    \draw [red,thick] (c) -- (cd);
    \draw [green,thick] (d) -- (dd);
    \draw [green,thick] (d) -- (bd);
    \draw [green,thick] (d) -- (de);
    \draw [green,thick] (d) -- (cd);
    \draw [orange,thick] (e) -- (ee);
    \draw [orange,thick] (e) -- (de);
\end{tikzpicture}}

@@

\begin{theorem}
    The independence polynomial~\cite{Harvey2018,Ferrin2014} of a graph $G = (V, E)$ can be computed in time $O(|V|^22^{{\rm tw}(G)})$.
\end{theorem}
\begin{proof}
    By contracting 
\end{proof}
Let $x_i = x$ and $w_i = 1$, \Cref{eq:idp} corresponds to the independence polynomial:
\begin{equation}\label{eq:idpdef}
I(G, x) = \sum_{k=0}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the number of independent sets of size $k$, $\alpha(G) \equiv \max_{I}\sum_{v\in I}w_v$ is the size of a maximum independent set and it is also called the independence number.
An independence polynomial is a useful graph characteristic related to, for example, the partition functions~\cite{Lee1952,Yang1952} and Euler characteristics of the independence complex~\cite{Bousquet2008, Levit2009}.
%Its connection to \Cref{eq:idp} can be understood as follows: the product over vertex tensor elements produces a factor $x^k$, where $k=\sum_i s_i$ counts the set size,
%and the product over edge tensor elements gives a factor $1$ for a configuration being in an independent set and $0$ otherwise. The summation counts the number of independent sets of size $k$. 
By assigning a real number to $x$, one can evaluate this independence polynomial for this specific value directly using tensor network contraction.
For example, the total number of independent sets can be evaluated as $I(G, 1)$.
However, instead of evaluating this polynomial for a certain value, we are more interested in knowing the coefficients of this polynomial, because this quantity tells us the counting of independent sets at different sizes.
%first elevate the tensor elements $0$s and $1$s in tensors $W(x)$ and $B$ from the standard number types to the additive identity, $\mymathbb{0}$, and multiplicative identity, $\mymathbb{1}$, of a commutative semiring as discussed in \Cref{sec:generic}.
%\purple{I'd
%say that you're about to describe this semiring, otherwise it seems like you're going to
%use the general definition of the semiring rather than a specific type of semiring}\blue{from $1$ to $\mymathbb{1}$ and $0$ to $\mymathbb{0}$ applies for a general semiring, I haven't change this part, let's discuss to make sure I understand your point.}\purple{Perhaps start the sentence with: Our semiring elements are a polynomial type...}
To this end, let us create a polynomial number data type by representing a polynomial $a_0 + a_1 x + \ldots + a_k x^k$ as a coefficient vector $a = (a_0, a_1, \ldots, a_k) \in \mathbb{R}^k$, e.g. $x$ is represented as $(0, 1)$.
Then we can define an algebra among coefficient vectors, including a redefinition of additive identity and multiplicative identity.
To avoid potential confusion, let us denote the additive identity as $\mymathbb{0}$, and the multiplicative identity is $\mymathbb{1}$.
The algebra between the polynomials number $a$ of order $k_a$ and $b$ of order $k_b$ is defined as
\begin{equation}
    \eqname{PN}
    \begin{split}
    a \oplus b &= (a_0 + b_0, a_1 + b_1, \ldots, a_{\max(k_a, k_b)} + b_{\max(k_a, k_b)}),\\
    a \odot b &= (a_0 + b_0, a_1b_0 + a_0b_1, a_{2}b_{0} + a_{1}b_{1} + a_{0}b_{2},  \ldots, a_{k_a} b_{k_b}),\\
    \mymathbb{0} &= (),  \\
    \mymathbb{1} &= (1), \label{eq:polynomial}
    \end{split}
\end{equation}
where $\oplus$ and $\odot$ are the standard polynomial addition and multiplication operations.
The multiplication can be evaluated in time $O(k\log(k))$ using the convolution theorem~\cite{Schonhage1971}, where $k=\max(k_a, k_b)$.
The tensors $W$ and $B$, which are introduced in the previous section (\Cref{sec:tnmap}), can thus be written as 
\begin{equation}
    W^{\rm PN} = \left(\begin{matrix}
        \mymathbb{1} \\
        (0,1)
    \end{matrix}\right),   
    \qquad \qquad

>>>>

\section{Tensor networks}\label{sec:tn}
A tensor network is composed of a collection of tensors $\mathcal{T}$, an indicator of which tensor is the output, and a map $\lambda$ from tensor to labels associated with tensor dimensions.
\begin{definition}[Tensor Network~\cite{Cirac2021, Orus2014}]
    A tensor network is a quadruple of $\mathcal{N} = (\Tensors, \Sigma, O \in \Tensors, \Inputs : \mathcal{T} \rightarrow \Sigma^k)$. $\Tensors$ is a set of tensors and $O$ is the output tensor. $\Sigma$ is a the alphabet of labels. $\Inputs$ is a map from a tensor to a vector of labels with the vector length the same as the tensor rank.
\end{definition}
The operation to evaluate a tensor network is called \textit{contraction}, which is defined as a summation of tensor element products over the labels not appearing in the output tensor.
For example, the matrix multiplication is a special tensor network
\begin{equation}
\mathcal{N}_{\rm matmul} = (\{A, B, C\}, C, T \mapsto \begin{cases} (i,j), T = A\\(j, k), T=B\\ (i, k), T=C\end{cases}, \{i,j,k\}),
\end{equation}
where $A, B$ and $C$ are matrices (two-dimensional tensors), among which $C$ is the output tensor, $(i,k),(i,j)$ and $(i,k)$ are labels associated with them. The contraction is defined as $C_{ik} = \sum_j A_{ij}B_{jk}$, where the subscripts are for tensor indexing, and the tensor dimensions with a same label must have the same size.
%can be viewed as a generalization of binary matrix multiplication to nary tensor contraction.
%Generalized Einstein's notation is often used to represent a tensor network, where we use a label in the subscript to represent a degree of freedom.
%One can enumerate these degrees of freedom and accumulate the product of tensor elements to the output tensor.
% Einstein's notation is often used to represent a tensor network, e.g.\ it represents the matrix multiplication between two matrices $A$ and $B$ as $C_{ik} = A_{ij}B_{jk}$,
% where we use a label in the subscript to represent a degree of freedom.
% One can enumerate these degrees of freedom and accumulate the product of tensor elements to the output tensor.
% In the standard Einstein's notation for tensor networks in physics, each index appears precisely twice, either both in input tensors (which will be summed over) or one in a input tensor and another in the output tensor.
% Hence a tensor network can be represented as a open simple graph,
% where a tensor is mapped to a vertex, a label shared by two input tensors is mapped to a normal edge and a label appears in the output tensor is mapped to an open edge.
% In this work, we adopt a more general form of tensor networks that do not restrict how many times a label can appear in tensors. 
The graphical representation of a tensor network is an open (meaning an edge can connect to external vertices) hypergraph, where an input tensor is mapped to a vertex and a label is mapped to a hyperedge that can connect an arbitrary number of vertices, while the labels appearing in the output tensor are open.
Our notation is a minor generalization of the standard tensor network notation used in physics as we do not restrict the number of times a label can appear in the tensors. 
While this generalized form is equivalent in representation power, it can have smaller contraction complexity as will be illustrated in \Cref{sec:tensorbad}.
% because one can easily translate a generalized tensor network to the standard notation by adding $\delta$ tensors, a high dimensional equivalence of the identity matrix.
%However, introducing $\delta$ tensors may significantly increase the complexity of the tensor contraction. We illustrate this subtle point in \Cref{sec:tensorbad}.

\begin{example}
\blue{maybe remove this example?}
\begin{equation}
\begin{split}
    \mathcal{T} &= \{A, B, V, C\},\\
    \Sigma &= \{i,j,k,l,m\},\\
    O &= C,\\
    \lambda &= T \mapsto \begin{cases}
    (j, k, m), &T = A\\
    (m, i, l), &T = B \\
    (j, m), &T = V \\
    (i, j, k), &T= C
    \end{cases}
    \end{split}
\end{equation}
is a tensor network that can be evaluated as $C_{ijk} = \sum_{ml}A_{jkm} B_{mil} V_{jm}$.
Its hypergraph representation is shown below, where we use different colors to represent different hyperedges.

@@

\section{Tensor network representation of independent sets} \label{sec:tnmap}
This section describes the reduction of the independent set problem to a tensor network contraction problem.
An alternative interpretation, perhaps more accessible to physicists, can be found in \Cref{sec:energymodel}, where we introduce the reduction from the energy model of hardcore lattice gases~\cite{Dyre2016, Fernandes2007}.
Let $G=(V,E)$ be an undirected graph with each vertex $v\in V$ being associated with a weight $w_v$.
An independent set $I \subseteq V$ is a set of vertices that for any vertex pair $u,v \in I$, we have $(u, v) \not\in E$.
To reduce the independent set problem on $G$ to a tensor network contraction, we first map each vertex $v\in V$ to a label $s_v \in \{0, 1\}$ of dimension $2$, where we use $0$ ($1$) to denote a vertex is absent (present) in $I$.
For each label $s_v$, we define a parameterized rank-one tensor associated with it as
\begin{equation}
    W^{(v)} = \left(\begin{matrix}
        1 \\
        x_v^{w_v}
    \end{matrix}\right),\label{eq:vertextensor}
\end{equation}
where $x_v$ is a variable associated with $v$.
Tensor elements can be indexed by $s_v$, e.g.\ $W^{(v)}_0=1$ is the first element associated with $s_v=0$ and $W^{(v)}_1=x_v^{w_v}$ is the second element associated with $s_v=1$.
Similarly, for each edge $(u, v) \in E$, we define a matrix $B$ indexed by $s_u$ and $s_v$ as
\begin{equation}
    \qquad \quad 
       B^{(u, v)} = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right), \label{eq:edgetensor}
\end{equation}
which encodes the independent set constraint since $B_{11} = 0$, meaning the two vertices cannot be both in the independent set. 
The corresponding tensor network $\mathcal{N}_{\rm IS}(G)$ is
\begin{equation}\label{eq:mistensornetwork}
\begin{split}
    \mathcal{T} &= \{W^{(v)} | v\in V\} \cup \{B^{(u, v)} | (u, v) \in E\} \cup \{P\}\\
    \Sigma &= \{s_v | v \in V\}\\
    O &= P\\
    \lambda &= T \mapsto \begin{cases}
                    (s_v), &T = W^{(v)}\\
                    (s_u, s_v), &T = B^{(u, v)}\\
                    (), & T = P
                    \end{cases}
    \end{split}
\end{equation}
where we use the superscripts to denote which vertex or edge a tensor is associated to.
$P$ is the output tensor of rank $0$, i.e. a scalar.
The contraction of this tensor network is
\begin{equation}\label{eq:idp}
    P(G) = \sum\limits_{s_1, s_2, \ldots, s_{|V|} = 0}^{1} \prod\limits_{v\in V} W^{(v)}_{s_v} \prod\limits_{(u,v) \in E} B^{(u,v)}_{s_u s_v},
\end{equation}
where the summation runs over all $2^{|V|}$ vertex configurations $\{s_1, s_{2}, \ldots,s_{|V|}\}$ and accumulates the product of tensor elements to the output $P$. 
% It is easy to see that an edge tensor element $B_{s_us_v}$ encodes the independent set constraint. It zeros out the product whenever both $u$ and $v$ are in the independent set.
A vertex tensor element $W^{(v)}_{s_v}$ contributes a multiplicative factor $x_v^{w_v}$ whenever $v$ is in the set.

@@

    \def\dx{5};
    \def\r{0.25cm}
    \foreach \x/\y/\e in {1.25/0/ab}
        \node[color=black,fill=black,dot=2.5*\r] at (\x+\dx,\y) (\e) {\scriptsize \color{white}{$B^{(a, b)}$}};
    \foreach \x/\y/\v in {0.3/0/a, 2.2/0/b}
        \node[color=black] at (\x+\dx,\y) (s\v) {$s_\v$};
    \foreach \x/\y/\v in {-0.5/0/a, 3.0/0.0/b}{
        \node[dot=\wr, color=black] at (\x+\dx,\y) (w\v) {};
        \node at (\x+\dx,\y+0.5) () {\scriptsize \color{black}{$W^{(v)}$}};
    }
    \draw [cyan,thick] (wa) -- (sa);
    \draw [cyan,thick] (sa) -- (ab);
    \draw [red,thick] (wb) -- (sb);
    \draw [red,thick] (sb) -- (ab);
\end{tikzpicture}}

In the graphical representation of the tensor network on the right panel,
we use a circle to represent a tensor, a hyperedge in cyan color to represent the degree of freedom $s_a$,
and a hyperedge in red color to represent the degree of freedom $s_b$.
Tensors sharing the same degree of freedom are connected by the same hyperedge.
The contraction of this tensor network has the following form:
\begin{equation}
    \begin{split}
    P({\rm K2}) &=
    \left(\begin{matrix} 1 & x_a^{w_a} \end{matrix}\right)
    \left(\begin{matrix} 1 & 1\\ 1  & 0 \end{matrix}\right)
    \left(\begin{matrix} 1 \\ x_b^{w_b} \end{matrix}\right)\\
    &= 1 + x_a^{w_a} + x_b^{w_b}.
    \end{split}
\end{equation}
The resulting polynomial represents 3 different independent sets $\{\}$, $\{a\}$, and $\{b\}$, with weights being $0$, $w_a$, and $w_b$ respectively.
\end{example}

For a general graph, it is computationally inefficient to evaluate \Cref{eq:idp} by directly summing up the $2^{|V|}$ products.
A better approach to evaluate a tensor network is: to find a good pair-wise tensor contraction order as a binary tree and then contract two tensors at a time by this order.
\begin{theorem}\label{thm:complexreal}
    If the addition and multiplication of tensor elements can be computed in time $O(1)$,
    the optimal time complexity for contracting the tensor network in \Cref{eq:mistensornetwork} for the independent set problem on graph $G = (V,E)$ is $O(|E|)2^{O({\rm tw}(G))}$.
\end{theorem}
\begin{proof}
Let us denote the line graph, a graph obtained by mapping an edge in the original graph to a vertex and connect two vertices if and only if their associated edges in the original original graph share a common vertex, of the hypergraph representation of a tensor network $\mathcal{N}$ as $L(\mathcal{N})$.
A contraction order of $\mathcal{N}$ corresponds to a tree decomposition of $L(\mathcal{N})$ and the largest intermediate tensor has a rank equal to the width of its tree decomposition~\cite{Markov2008}.
Therefore, an optimal (in terms of space complexity) contraction order corresponds to the tree decomposition of $L(\mathcal{N})$ with the smallest width (or the treewidth ${\rm tw}(L(\mathcal{N})$). The contraction complexity is $O(|\mathcal{N}|)2^{O({\rm tw}(L(\mathcal{N})))}$, where $|\mathcal{N}|$ is the number of tensors in $\mathcal{N}$.
For the independent set problem on graph $G = (V, E)$, the line graph of the hypergraph representation of the tensor network in \Cref{eq:mistensornetwork} is the same as the graph $G$ up to some isolated vertices, hence the contraction complexity of this tensor network with an optimal contraction order is $O(|E|)2^{O({\rm tw}(G))}$.
\end{proof}

In practice, it is difficult to find an optimal contraction order for large tensor networks because finding the treewidth is a well-known NP-hard problem.
However, it is easy to find a close-to-optimal contraction order within typically a few minutes with a heuristic algorithm~\cite{Kourtis2019, Kalachev2021}.
%The pairwise tensor contraction also makes it possible to utilize basic linear algebra subprograms (BLAS) functions to speed up the computation for certain tensor element types.
For large-scale applications, it is also possible to slice over certain degrees of freedom to reduce the space complexity, i.e.\
loop and accumulate over certain degrees of freedom so that one can have a smaller tensor network inside the loop due to the removal of these degrees of freedom.

\begin{example}\label{eg:tensorcontraction}
\blue{maybe remove this example too?}
In this example, we map a 5-vertex graph (left) to a tensor network (right) and show how the contraction order reduces the time and space complexities.

\centerline{\begin{tikzpicture}[
dot/.style = {circle, fill, minimum size=#1,
            inner sep=0pt, outer sep=0pt},
dot/.default = 6pt  % size of the circle diameter 
                ]  
    \def\dx{0};
    \def\r{0.25cm}
    \filldraw[fill=black] (\dx,0) circle [radius=\r];
    \filldraw[fill=black] (\dx,1.5) circle [radius=\r];
    \filldraw[fill=black] (1.5+\dx,0) circle [radius=\r];
    \filldraw[fill=black] (1.5+\dx,1.5) circle [radius=\r];
    \filldraw[fill=black] (2.5+\dx,2.5) circle [radius=\r];
    \draw [black,thick] (\dx,0) -- (\dx,1.5);
    \draw [black,thick] (\dx,0) -- (1.5+\dx,0);
    \draw [black,thick] (\dx,1.5) -- (1.5+\dx,1.5);
    \draw [black,thick] (1.5+\dx,0) -- (1.5+\dx,1.5);
    \draw [black,thick] (1.5+\dx,0) -- (\dx,1.5);
    \draw [black,thick] (2.5+\dx,2.5) -- (1.5+\dx,1.5);
    \node[color=white] at (\dx,0) {a};
    \node[color=white] at (\dx,1.5) {b};
    \node[color=white] at (1.5+\dx,0) {c};
    \node[color=white] at (1.5+\dx,1.5) {d};
    \node[color=white] at (2.5+\dx,2.5) {e};
    \def\dx{5};
    \def\r{0.25cm}
    \foreach \x/\y/\u/\v in {0.75/0/a/c, 0/0.75/a/b, 1.5/0.75/c/d, 0.75/1.5/b/d, 0.75/0.75/b/c, 2/2/d/e}
        \node[color=white,fill=black,dot=2.5*\r] at (\x+\dx,\y) (\u\v) {\scriptsize $B^{(\u,\v)}$};
    \foreach \x/\y/\v in {0/0/a, 0/1.5/b, 1.5/0/c, 1.5/1.5/d, 2.5/2.5/e}
        \node[color=black] at (\x+\dx,\y) (\v) {$s_\v$};
    \foreach \x/\y/\v in {-0.5/-0.5/a, -0.5/2.0/b, 2.0/-0.5/c, 2.0/1.0/d, 3.0/2.0/e}
        \node[color=white,fill=black,dot=\r] at (\x+\dx,\y) (\v\v) {};
    \foreach \x/\y/\v in {-0.5/-0.5/a, -0.5/2.0/b, 2.0/-0.5/c, 2.0/1.0/d, 3.0/2.0/e}
        \node[color=black] at (\x+\dx+0.6,\y) {\scriptsize $W^{(\v)}$};
    \draw [cyan,thick] (a) -- (aa);
    \draw [cyan,thick] (a) -- (ab);
    \draw [cyan,thick] (a) -- (ac);
    \draw [blue,thick] (b) -- (bb);
    \draw [blue,thick] (b) -- (ab);
    \draw [blue,thick] (b) -- (bc);
    \draw [blue,thick] (b) -- (bd);
    \draw [red,thick] (c) -- (cc);
    \draw [red,thick] (c) -- (ac);
    \draw [red,thick] (c) -- (bc);
    \draw [red,thick] (c) -- (cd);
    \draw [green,thick] (d) -- (dd);
    \draw [green,thick] (d) -- (bd);
    \draw [green,thick] (d) -- (de);
    \draw [green,thick] (d) -- (cd);
    \draw [orange,thick] (e) -- (ee);
    \draw [orange,thick] (e) -- (de);
\end{tikzpicture}}

@@

Let $x_i = x$ and $w_i = 1$, \Cref{eq:idp} corresponds to the independence polynomial:
\begin{equation}\label{eq:idpdef}
I(G, x) = \sum_{k=0}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the number of independent sets of size $k$, $\alpha(G) \equiv \max_{I}\sum_{v\in I}w_v$ is the size of a maximum independent set and it is also called the independence number.
An independence polynomial is a useful graph characteristic related to, for example, the partition functions~\cite{Lee1952,Yang1952} and Euler characteristics of the independence complex~\cite{Bousquet2008, Levit2009}.
%Its connection to \Cref{eq:idp} can be understood as follows: the product over vertex tensor elements produces a factor $x^k$, where $k=\sum_i s_i$ counts the set size,
%and the product over edge tensor elements gives a factor $1$ for a configuration being in an independent set and $0$ otherwise. The summation counts the number of independent sets of size $k$. 
By assigning a real number to $x$, one can evaluate this independence polynomial for this specific value directly using tensor network contraction.
For example, the total number of independent sets can be evaluated as $I(G, 1)$.
However, instead of evaluating this polynomial for a certain value, we are more interested in knowing the coefficients of this polynomial, because this quantity tells us the counting of independent sets at different sizes.
%first elevate the tensor elements $0$s and $1$s in tensors $W(x)$ and $B$ from the standard number types to the additive identity, $\mymathbb{0}$, and multiplicative identity, $\mymathbb{1}$, of a commutative semiring as discussed in \Cref{sec:generic}.
%\purple{I'd
%say that you're about to describe this semiring, otherwise it seems like you're going to
%use the general definition of the semiring rather than a specific type of semiring}\blue{from $1$ to $\mymathbb{1}$ and $0$ to $\mymathbb{0}$ applies for a general semiring, I haven't change this part, let's discuss to make sure I understand your point.}\purple{Perhaps start the sentence with: Our semiring elements are a polynomial type...}
To this end, let us create a polynomial number data type by representing a polynomial $a_0 + a_1 x + \ldots + a_k x^k$ as a coefficient vector $a = (a_0, a_1, \ldots, a_k) \in \mathbb{R}^k$, e.g. $x$ is represented as $(0, 1)$.
Then we can define an algebra among coefficient vectors, including a redefinition of additive identity and multiplicative identity.
To avoid potential confusion, let us denote the additive identity as $\mymathbb{0}$, and the multiplicative identity is $\mymathbb{1}$.
The algebra between the polynomials number $a$ of order $k_a$ and $b$ of order $k_b$ is defined as
\begin{equation}
    \eqname{PN}
    \begin{split}
    a \oplus b &= (a_0 + b_0, a_1 + b_1, \ldots, a_{\max(k_a, k_b)} + b_{\max(k_a, k_b)}),\\
    a \odot b &= (a_0 + b_0, a_1b_0 + a_0b_1, a_{2}b_{0} + a_{1}b_{1} + a_{0}b_{2},  \ldots, a_{k_a} b_{k_b}),\\
    \mymathbb{0} &= (),  \\
    \mymathbb{1} &= (1), \label{eq:polynomial}
    \end{split}
\end{equation}
where $\oplus$ and $\odot$ are the standard polynomial addition and multiplication operations.
